Phylogenetics and the human microbiome.

### 2014-06-24 response/response.txt

> P12 “Several groups are continually revising taxonomy...” and P13 “An
> obvious application of phylogenetics is to perform taxonomic
> classification...” We have here two long and technical paragraphs
> (e.g. F-measure, precision and recall, but what is classified and
> how?). You should check this again for clarity and to ensure
> self-contentedness.

You are right. I've decided to just eliminate the technical sentences.

> P16 connecting taxa in the set (Faith 1992).

I have changed this to:

"The (unweighted and unrooted) PD of a set of taxa $S$ is simply the total of the lengths of the branches which connect taxa in $S$ (Faith 1992)."

@@

> Fig 1 The "graphical formulae" [l(segments) A(segments)] at the bottom
> of the figure do not help. This can be better described by
> checking-improving the figure legend, notably "the abundance of the
> taxa on either side of the branch" (how is it computed? give an
> example if there are several options)

I have removed the graphical formulae, and given an example abundance-weighted measure.

@@

> Fig 2 Your two cases are extreme. Having one (or a few) shared (brown)
> edges in the right tree, and one unshared (blue or green branch) edge
> in the left tree would help. See also comments of referee 2.

I have made it less extreme according to your suggestion, thanks.

@@

> Fig 3 and its legend, as well as the corresponding text P17-18 are
> quite unclear to me.... It would be interesting to have more
> explanations, both on the math side and on the application to sequence
> reads.

I feel that by expanding the explanation here I would contribute to the un-evenness of coverage described by the other reviewers. The earth-movers distance is a standard way to compare probability distributions on metric spaces and has a Wikipedia page. I have added a classical and a recent book reference.

@@

Reviewer: 2
-----------

> As noted in my first review, I like the idea of this manuscript,and
> much of the content.  The revised version is very close to the first
> version, so I assume the author is happy with it.  I certainly
> appreciated the diff.pdf.

> I am still not clear how the author chooses what needs technical
> discussion and what does not, and more technical detail has been added
> to this version.  However other, perhaps more technically-inclined
> reviewers, did not really pick this up, so, in the interests of
> efficiency, I won't belabour the point. The author must have a clear
> view who the intended audience is.

Well, I chose to discuss things at depth those subjects I thought would
interest the phylogenetics community. I have removed discussion of one
outlier as described above, but am open to hearing about others that
should be removed.

@@

> I will say that, while figure 2 does a fairly good job of illustrating
> the critical UniFrac method (though the common brown internal edges
> could be named explicitly in the caption), the definition in the text
> does not: "The name UniFrac is a contraction of “Unique Fraction,”
> which refers to the fact that the original UniFrac definition compares
> the fraction of edges that connect only tips from one sample via the
> shortest path rather than edges that connect between samples (Fig. 2;
> Lozupone and Knight 2005)" makes very little sense to me, particularly
> "edges between samples". Could this be reworded in a more intuitive
> way that refers explicitly to the tree(s) being inferred and the tips
> used to infer the trees?  So, for instance, does one infer one tree
> per sample and then one tree of all samples?

Thank you for your comments-- the text has been updated (I discussed
these changes via email with Arne directly).

### 2014-06-24 sbreview.tex
Addressing the last of the comments.

\newcommand{\pdLegend}{
    Unweighted phylogenetic diversity (PD, left) and an abundance-weighted PD measure (right), where taxa present in a sample are shown as circles and abundances are shown as the size of the circles.
    Unweighted PD takes the sum of the branch lengths $\ell$ connecting samples.
    Abundance-weighted measures take a weighted sum of branch lengths where weight is determined in some way by the abundance of the taxa on either side of the branch: if we give edges width according to their weight, the abundance-weighted measure can be thought of as the sum of the areas $A$ of edges.
}
\forarxiv{
\begin{wrapfigure}[25]{r}{2.2in}
  % \capstart
  \vspace{-14pt}
  \begin{center}
    \includegraphics[width=2.2in]{figures/pd.pdf}
  \end{center}
  \vspace{-10pt}
  \caption{\pdLegend}
  \label{fig:pd}
\end{wrapfigure}
}

@@

The name UniFrac is a contraction of ``Unique Fraction,'' which refers to the fact that the original UniFrac definition compares the fraction of edges that connect only tips from one sample via the shortest path rather than edges that connect between samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'', defined as mass times distance, required to move probability mass in one distribution to another along the tree.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \citet{faith2009cladistic}, who point out related measures in Faith's earlier work.

@@

\newcommand{\dirtpilesLegend}{
    Part of a minimal mass movement to calculate the earth-mover's distance between two probability distributions on a phylogenetic tree.
    For this, each probability distribution is considered as a configuration of dirt piles on the tree, and the distance between two such dirt pile configurations is defined to be the minimum amount of physical ``work'' required to move the dirt in one configuration to the other.
}
\forarxiv{
\begin{wrapfigure}[24]{l}{1.5in}
  % \capstart
  \vspace{-17pt}
  \begin{center}
    \includegraphics[width=1.2in]{figures/tree_dirtpiles.pdf}
  \end{center}
  \vspace{-5pt}
  \caption{\dirtpilesLegend}
  \label{fig:dirtpiles}
\end{wrapfigure}
}

>>>>

\newcommand{\pdLegend}{
    Unweighted phylogenetic diversity (PD, left) and an abundance-weighted PD measure (right), where taxa present in a sample are shown as circles and abundances are shown as the size of the circles.
    Unweighted PD takes the sum of the branch lengths connecting samples.
    Abundance-weighted measures take a weighted sum of branch lengths where weight is determined in some way by the abundance of the taxa on either side of the branch: if we give edges width according to their weight, the abundance-weighted measure can be thought of as the sum of the total area of the edges.
    One such abundance-weighted measure simply takes the absolute value of the difference of the total taxon abundance on one side compared to the other.
}
\forarxiv{
\begin{wrapfigure}[25]{r}{2.2in}
  % \capstart
  \vspace{-14pt}
  \begin{center}
    \includegraphics[width=2.2in]{figures/pd.pdf}
  \end{center}
  \vspace{-10pt}
  \caption{\pdLegend}
  \label{fig:pd}
\end{wrapfigure}
}

@@

To calculate the UniFrac divergence between two samples of reads, one builds a tree on both samples and then performs a calculation on the branch lengths according whether branches have descendants in both samples or one sample only.
Specifically, the original UniFrac divergence is the fraction of the total branch length of the tree coming from branches with descendant leaves in exactly one of the two samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance \citep{monge1781memoire,Villani2003-wv} between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'', defined as mass times distance, required to move probability mass in one distribution to another along the tree.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \citet{faith2009cladistic}, who point out related measures in Faith's earlier work.

@@

\newcommand{\dirtpilesLegend}{
    Part of a minimal mass movement to calculate the earth-mover's distance between two probability distributions on a phylogenetic tree.
    For this, each probability distribution is considered as a configuration of dirt piles (round bumps in the figure) on the tree, and the distance between two such dirt pile configurations is defined to be the minimum amount of physical ``work'' required to move the dirt in one configuration to the other.
}
\forarxiv{
\begin{wrapfigure}[24]{l}{1.5in}
  % \capstart
  \vspace{-17pt}
  \begin{center}
    \includegraphics[width=1.2in]{figures/tree_dirtpiles.pdf}
  \end{center}
  \vspace{-5pt}
  \caption{\dirtpilesLegend}
  \label{fig:dirtpiles}
\end{wrapfigure}
}
